---
title: "Assignment 5.2 Washington Donations"
author: "Larry Heckel"
date: "April 13, 2019"
output: pdf_document
---

# Assignment

As an aspiring Data Scientist, you have heard that a majority of the time you can expect to spend on data acquisition and cleanup, aka data wrangling. For this assignment, you will be working with some data that has been manually entered into Donations tracking documents for Washington D.C. and Nebraska.   

1. How many variables does the data set have, explain any peculiarities of the source data such as how     missing values are discovered.

2. Data importing and cleaning steps are explained in the text and in the DataCamp exercises (tell me     why you are doing the data cleaning activities that you perform) and follow a logical process.

3. Once your data is clean, show what the ﬁnal data set looks like. However, do not print off a data      frame with 200+ rows; show me the data in the most condensed form possible.

4. Add a variable to your data set called Average Donation Amount. Calculate the Average Donation         Amount per transaction and insert that value into a new data set.

5. Provide summary information using the appropriate R function(s) about the Donation Amount variable     in your cleaned data set. Provide a written summary paragraph that describes your findings.

```{r echo=TRUE}
# libraries to be used
library(ggplot2)
library(readr)
library(dplyr)
library(stringr)
library(psych)

# read the files individually
washDonation <- read_csv("donations-2015-2016-DC.txt")
nebrDonation <- read_csv("donations-2015-2016-NE.txt")
# View(washDonation)
# str(washDonation)
# View(nebrDonation)
# str(nebrDonation)

#Combine the two data frames into one
comboDonation <- rbind(washDonation, nebrDonation)
View(comboDonation)
str(comboDonation)

```
```{r echo=FALSE}
################

options(scipen=100)
options(digits=2)
```

## Section 1

How many variables does the data set have, explain any peculiarities of the source data such as how   missing values are discovered.

  The data set contains 16 variables. Each row, or observation, is the record of a donation to a candidate for a political office. In each row, there is one variable, called "donation_amount", which records the dollar amount of the donation. The other 15 variables contain data about the donation, such as candidate and donor information.
  
The next step is to determine if there are missing values in any of the observations.

```{r echo=TRUE}
# create data frame with rows of data that have missing values 
comboNA <- comboDonation[!complete.cases(comboDonation),]
comboNA
View(comboNA)
#how many missing are in each column/variable?
colSums(is.na(comboNA))

```  
The two most important variables in the data set are the candidate name and the donation amount.
Neither of those variables has any missing values, so we will be able to answer our most basics metrics questions, which are:

1. How many donations did each candidate receive:
2. What is the total amount of money each candidate received?
3. What is the average dollar amount of each donation, for each candidate?

Additionally, the following columns also do not have any missing values:

1. Party Affiliation
2. Year of Election
3. Candidate State
4. Candidate Office
5. Committee Name
6. Donor State
7. Donation Date

The below columns have missing values (Number missing of 112,191 total donations made).

1. Incumbent/Challenger Status (8 missing)
2. Committee Party (8,222 missing)
3. Donor Name (9 missing)
4. Donor City (4 missing)
5. Donor Zip (19 missing)
6. Donor Employer (3,453 missing)
7. Donor Occupation (2,572 missing)



## Section 2

Data importing and cleaning steps are explained in the text and in the DataCamp exercises (tell me    why you are doing the data cleaning activities that you perform) and follow a logical process.

1. First, I will deal with missing metadata values. Some of the missing metatdata is not important
    for the analysis at hand, only that it is not empty, so for those variables, I will provide a
    value that will denote that it was not originally provided. For the variables that are most
    likely required for a good analysis, I will derive what I feel are the correct values and provide
    them. The details for each variable are listed below, and the code executed on the data frame
    below that.

    The below columns have missing values (Number missing of 112,191 total donations made).

    a. Incumbent/Challenger Status (8 missing)
        The two candidates for whom this value is missing only have these 8 rows, and don't have any
          other entries in the data set. Therefore, we are not able to derive an appropriate value, 
          so I will fill with "UNK" for UNKNOWN.
    b. Committee Party (8,222 missing)
        Is the same as Party Affiliation, so will fill with that value
    c. Donor Name (9 missing)
        Not an important variable, will fill with "Not Given".
    d. Donor City (4 missing)
        Not an important variable, will fill with "Not Given"
    e. Donor Zip (19 missing)
        Not an important variable, will fill with "99999"
    f. Donor Employer (3,453 missing)
         Not an important variable, will fill with "Not Given"
    g. Donor Occupation (2,572 missing) 
        Not an important variable, will fill with "Not Given"

```{r echo=TRUE}
#Create new data frame called comboClean, for the clean metadata
#Start from comboDonation with the first data mutation, then each subsequent
#mutation is on comboClean.
comboClean <- comboDonation %>% mutate(donor_zip=ifelse(is.na(donor_zip), 99999, donor_zip))
comboClean <- comboClean %>% mutate(donor_name=ifelse(is.na(donor_name), "Not Given", donor_name))
comboClean <- comboClean %>% mutate(donor_city=ifelse(is.na(donor_city), "Not Given", donor_city))
comboClean <- comboClean %>% mutate(donor_employer=ifelse(is.na(donor_employer), "Not Given", donor_employer))
comboClean <- comboClean %>% mutate(donor_occupation=ifelse(is.na(donor_occupation), "Not Given", donor_occupation))
comboClean <- comboClean %>% mutate(committee_party=ifelse(is.na(committee_party), party_affiliation, committee_party))

#check for the candidates who don't have an incumber/challenger flag set
mandel <- comboDonation %>%
  filter(str_detect(candidate_name, "MANDE"))
lindbeck <- comboDonation %>%
  filter(str_detect(candidate_name, "LINDB"))
View(mandel)
View(lindbeck)

comboClean <- comboClean %>% mutate(incumbent_challenger_status=ifelse(is.na(incumbent_challenger_status), "UNK", incumbent_challenger_status))

#now check to ensure that there are no NA values
colSums(is.na(comboClean))

```
2. Now that all of the metadata is clean, we will look at the actual donation values, to see what the
    data distribution looks like, and if there are outliers. 
    
      a. I will start with by checking for values less than or equal to 0 (0 or negative), because
          those would not make sense in terms of donations. In the context of our understanding,
          donations must be dollars given to a candidate, not from the candidate to the donor. We
          will remove those rows from the data set. There are 490 observations that meet this
          criteria, and are removed.
```{r echo=TRUE}
donCount <- count(comboClean, vars=donation_amount <= 0)
str(donCount)

comboCleanDonation <- subset(comboClean,subset = (donation_amount > 0))
str(comboCleanDonation)

```
      b. Next I'll check the data distribution of the donation amounts, to see if there is anything
          out of the ordinary with it. I'll run summary statistics and some data distribution graphs
          to also see the data visually. See the results below, and conclusions that follow.
          
```{r echo=TRUE}
summary(comboCleanDonation$donation_amount)
```
```{r echo=FALSE}
boxplot_comboCleanDonation <- ggplot(data=comboCleanDonation,aes(y=donation_amount)) +
  geom_boxplot() 
boxplot_comboCleanDonation

hist_comboCleanDonation_Norm <- ggplot(data=comboCleanDonation,aes(x=donation_amount)) + 
  geom_histogram(aes(y=..count.., fill=..count..)) +
  scale_x_log10("donation_amount") +
  ggtitle("Donation Amount Histogram") +
  ylab("Number of Donations") 
hist_comboCleanDonation_Norm

donationHigh <- subset(comboCleanDonation, comboCleanDonation$donation_amount > 10000)
str(donationHigh)
```
          The distribution of the donation amounts is heavily skewed to the left, with the median
          being 100, meaning that half of the donations are $100 or less. Additionally, the mean is
          $399, so there are fewer, exceptionally larger donations in the data set.
          
          The range and distribution of the amounts makes visualizing the data difficult to
          represent, as can be seen from the box plot graph. To be able to view the amounts, 
          I have scaled the data to the log10 of the donation amount. We can still see
          in this graph the left skew of the data, and how a small number of exceptionally high
          donations affect the view.
          
          So I looked at the donations of amounts greater than $10,000, to see if they make sense,
          and they do, so I will keep them in the data set. The Donor Occupations indicate that the
          Donors are of sufficient means to make these types of large donations, so they stay.
          
          There are no other data cleansing tasks to do, so I will proceed to the analysis.

## Section 3

Once your data is clean, show what the ﬁnal data set looks like. However, do not print off a data     frame with 200+ rows; show me the data in the most condensed form possible.

```{r echo=TRUE}
str(comboCleanDonation)
```
## Section 4

Add a variable to your data set called Average Donation Amount. Calculate the Average Donation        Amount per transaction and insert that value into a new data set.

```{r echo=TRUE}
#calculate the Average Donation Amount,
#Create vector to append to data frame
AvgDonAmt <- c(mean(comboCleanDonation$donation_amount))
names(AvgDonAmt) <- c("Average Donation Amount")

#Append the Average to each row in the data frame,
#show the structure, to confirm success
comboCleanDonationAvg <- cbind(comboCleanDonation, AvgDonAmt)
str(comboCleanDonationAvg)


```
## Section 5

Provide summary information using the appropriate R function(s) about the Donation Amount variable    in your cleaned data set. Provide a written summary paragraph that describes your findings.

```{r echo=TRUE}
options(scipen=100)
options(digits=2)
summary(comboCleanDonation$donation_amount)
describe(comboCleanDonationAvg$donation_amount)
```
### Conclusion

As I had noted above, The distribution of the donation amounts is heavily skewed to the left, with
  the median being 100, meaning that half of the donations are $100 or less. Additionally, the mean
  is $399, so there are fewer, exceptionally larger donations in the data set. Finally, the 3rd
  Quantile value is only $500, meaning that 75% of the donations are that amount or less.
  
Further summary statistics confirm all of this, as the skew and kurtosis values both indicate the
  left skew and high peak of the graph.
  
What all this means in real terms is that the candidates have a large number of donors who give
  smaller donations, and a few who give much larger donations. This is consistent with other reports
  of campaign contributions in other election cycles.
  
Although out of scope of this analysis, the data set is cleaned and able to provide answers to more
  insightful questions, such as those cited at the beginning of this write-up, and others, such as
  total and average donations by political party, by election race, state, office, and all of the
  metadata variables in the data set. One additional note is that, if these analysis were foreseen,
  then we would go back and spend much more effort cleaning the data, looking to derive useful
  values, rather than simply filling them in with "UNK" or "NOT PROVIDED".
  

